---
title: "anthroDisturbance_Generator Manual"
subtitle: "v.`r SpaDES.core::moduleMetadata(module = 'anthroDisturbance_Generator', path = '..')$version`"
date: "Last updated: `r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 4
    html_preview: false
editor_options:
  chunk_output_type: console
---

# anthroDisturbance_Generator Module

```{r setup-anthroDisturbance_Generator, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, results = "hold") ## change to eval = TRUE if all chunks are to be evaluated
```

#### Authors:

`r paste(as.character(SpaDES.core::moduleMetadata(module = "anthroDisturbance_Generator", path = "..")$authors), sep = ", ")`

## Module Overview

### Module summary

This module generates new anthropogenic disturbances forward in time. Given the
`disturbanceParameters` table and a `disturbanceList` holding both current and
potential disturbances (as produced by `anthroDisturbance_DataPrep` and,
optionally, `potentialResourcesYT_DataPrep`), it grows, generates, and connects
disturbances year by year according to each type's rate and size.

The module was originally developed for the Northwest Territories (its built-in
default sample data is the union of BCR6 and NT1), but the structure is general.
This is the FOR-CAST fork used in the Yukon Northern Mountain Caribou project,
where the pipeline supplies Yukon inputs.

### Module inputs and parameters

The two structuring inputs are `disturbanceList` (from a DataPrep / potential
resources module) and the `disturbanceParameters` `data.table`, whose columns
are:

- `dataName`: sector grouping (e.g., Energy, Settlements, OilGas, Mining,
  Forestry, Roads);
- `dataClass`: the specific class, always prefixed `potential` (e.g.,
  potentialSettlements, potentialWindTurbines, potentialCutblocks);
- `disturbanceType`: one of three general types --
    1. Enlarging (e.g., potentialSettlements, potentialSeismicLines): the
       potential layer equals the current layer and is only buffered over time;
    2. Generating (e.g., potentialWindTurbines, potentialOilGas,
       potentialMineral, potentialForestry): the potential layer marks where
       structures may appear based on a rate;
    3. Connecting (e.g., potentialPipelines, potentialTransmission,
       potentialRoads): the potential layer needs the current transmission,
       pipeline, and road network, and depends on what Generating produces;
- `disturbanceRate`: yearly generation rate for Enlarging and Generating types
  (NA for Connecting). If missing when needed, the module derives it from data
  (e.g., ECCC), falling back to a yearly average of 0.2% of current disturbance;
- `disturbanceSize`: target size in m2 for Generating types (NA for Enlarging /
  Connecting); derived from data if not supplied;
- `disturbanceOrigin`: the `dataClass` used as the "original" to modify
  (Enlarging, Generating) or as the origin point for Connecting types;
- `disturbanceEnd`: end points for Connecting layers (e.g., new wind turbines
  connect into power lines/roads; new oilGas into pipelines/roads);
- `disturbanceInterval`: interval at which the disturbance recurs.

Other inputs include the study area (`studyArea`), matching raster
(`rasterToMatch`), current burn raster (`rstCurrentBurn`), `disturbanceDT`, and
optional `DEM` / `featuresToAvoid` layers used to constrain linear-feature
placement.

The full list of module inputs:

```{r moduleInputs-anthroDisturbance_Generator, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_inputs <- SpaDES.core::moduleInputs("anthroDisturbance_Generator", "..")
knitr::kable(df_inputs)
```

This module exposes many parameters controlling generation, clustering, and
data-derived rate calculation. The complete, current list (with defaults and
descriptions) is generated directly from the module metadata:

```{r moduleParams-anthroDisturbance_Generator, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_params <- SpaDES.core::moduleParams("anthroDisturbance_Generator", "..")
knitr::kable(df_params)
```

Note `runClusteringInParallel` (default `FALSE`) enables parallel line
clustering; when enabled, `maxClusterCores` caps the number of workers so the
module cannot over-subscribe cores when nested under `targets` branching.

### Events

The core events are:

1. `calculatingSize`: if disturbance sizes are not supplied in
   `disturbanceParameters`, uses current disturbance data to estimate the mean
   and variation of each type's size, later drawn from a normal distribution;
2. `calculatingRate`: if disturbance growth rates are not supplied, derives them
   -- using the 2010 and 2015 human-footprint datasets to compute a 5-year
   growth rate per type -- and applies the result in `generatingDisturbances`;
3. `generatingDisturbances`: the core event (`generateDisturbances()`). It sizes
   the study area, masks current disturbances, then (a) grows Enlarging types by
   buffering existing features, (b) generates Generating types by iteratively
   selecting the highest-potential areas (from the potential rasters) until the
   target rate is met, and (c) connects new features to linear networks (roads,
   power lines). It rasterizes and merges results so already-chosen pixels are
   unavailable next iteration. Reschedules every `runInterval`;
4. `updatingDisturbanceList`: incorporates the new roads/power lines, adds the
   generated disturbances to the current ones, and replaces enlarged settlements
   and seismic lines with their buffered versions. Reschedules every `runInterval`.

`init` validates inputs and schedules the events; a `plot` event runs on the
`.plotInitialTime`/`.plotInterval` schedule.

### Plotting

The `plot` event routes a disturbance-presence map through `SpaDES.core::Plots()`.
It rasterizes the current disturbance layers to a presence raster over the study
area and draws it with `tidyterra`/`ggplot2` (viridis fill with the study-area
outline), writing the figure(s) to the module's figures path and registering
them as outputs.

### Saving

When `saveInitialDisturbances = TRUE` and/or `saveCurrentDisturbances = TRUE`,
per-step disturbance rasters are written to `Paths[["outputPath"]]` (prefixed by
sector name and suffixed by year) and registered via
`SpaDES.core::registerOutputs()` so `targets` tracks them.

### Module outputs

```{r moduleOutputs-anthroDisturbance_Generator, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}
df_outputs <- SpaDES.core::moduleOutputs("anthroDisturbance_Generator", "..")
knitr::kable(df_outputs)
```

`disturbanceList` is the updated list of current + potential disturbances;
`currentDisturbanceLayer` is a per-year list of rasters of all current
disturbances, used to filter pixels already disturbed.

### Links to other modules

Part of the anthropogenic-disturbance module collection, run after
`anthroDisturbance_DataPrep` and (optionally) `potentialResourcesYT_DataPrep`.
`rstCurrentBurn` is typically supplied by a fire module (e.g., `scfm`). The
collection combines with landscape-simulation modules (e.g., `Biomass_core`) and
caribou modules to improve realism in forecasts.

### Getting help

- <https://github.com/FOR-CAST/anthroDisturbance_Generator/issues>
